Fast Hyperparameter Selection for Graph Kernels via Subsampling and\\Multiple Kernel Learning

### 2016-11-26 paper.tex
F

With kernel methods, a user has usually to deal with several of hyperparameters to fix manually. The optimization of these hyperparameters can strongly affect the predictive performance of the method.
Usually, these parameters are: (i) the choice of the kernel to adopt; (ii) the hyperparameters of the selected kernel; (iii) the hyperparameters of the learning algorithm.
A commonly adopted and simple approach to hyperparameters optimization is to fix a-priori some possible values for each parameter (the combinations of these parameters are referred to as \textit{parameters grid}).
Then, the learning algorithm is trained for each possible parameters combination, and its performances are evaluated. This approach is commonly referred to as grid-search.

With this approach, a learning procedure has to be performed for each hyperparameter combination, possibly making it very time consuming.
A possible alleviation to this problem has been proposed in~\cite{FabioAiolliMicheleDonini2015,Massimo2016}, where Multiple Kernel Learning (MKL) has been used for combining different kernel matrices in a single learning procedure, instead of selecting the best parameters configuration.
However, in the case of kernel algorithms, a different kernel matrix has to be computed for each kernel and for each of its hyperparameters values in the grid. This step can also be time consuming, especially when dealing with data in a structured form.

In this paper, we propose a new approach to parameter selection that, based on sub-sampling and MKL, is able to speedup the parameter selection phase by orders of magnitude on real-world datasets.

@@

In~\cite{Costa2010} the Neighborhood Subgraph Pairwise Distance Kernel (NSPDK) is presented. This kernel computes the exact matches between pairs of subgraphs with controlled size and distance (hyperparameters $r$ and $d$, respectively). 
%The kernel has $O(|V| |V_h| |E_h| \log |E_h|)$ time complexity, where $|V_h|$ and $|E_h|$ are the number of nodes and the number of edges of the subgraph obtained by a breadth-fist visit of depth $h$. 
% The complexity of the kernel is $O(|V| |V_h| |E_h| \log |E_h|)$, where $|V_h|$ and $|E_h|$ are the number of nodes and the number of edges of the subgraph obtained by a breadth-fist visit of depth $h$. 
%The authors state that, for small values of the subgraph size and distance, the complexity of the kernel becomes in practice linear.

@@

\section{Experimental Assessment}
We conducted our experiments on several real-world binary datasets, that are MUTAG, CPDB, GDD, AIDS, NCI1, NCI109, CAS (see \cite{DaSanMartino2016,Shervashidze2011} for a more detailed description).
%AIDS~\cite{Weislow1989a}, CAS\footnote{http://cheminformatics.org/datasets/bursi/},
%CPDB~\cite{Helma2004a}, NCI1~\cite{Wale2006} and GDD~\cite{dobson2003}.
%The first four datasets contain chemical particles encoded in graph form, while the latter encodes proteins.
%{\color{red} The AIDS dataset contains chemical compounds (encoded in graph form), labeled according to their antiviral activity; CAS and CPDB store compounds labeled according to their mutagenicity; NCI1 and NCI109 consist of chemical compounds screened for activity against non-small lung cancer cells; GDD is composed of X-ray crystal structures of proteins; MUTAG contains %MUTAG serve per far vedere che funziona bene anche su dataset piccolissimi
%[...].
%All the nodes are labeled and there are no self loops.}
%
The kernels used in our experiments are:  NSPDK  with $h \in \{0,1\dots9\}$ and $d \in \{0,1\dots9\}$,  ODD$_{ST}$ with $h \in \{0,1\dots9\}$ and $\lambda \in \{0.5, 0.7,0.8\dots1.2,1.4,1.6,1.8\}$, WL with $h \in \{0,1\dots9\}$.
In all models, we considered kernels calculated in less that 1 hour, rejecting others.
The baselines used to compare our method consist on a SVM with kernels (a)\kval: the kernel is the one which has the best validation score, (b)\ksum: the kernel is the average of all base kernels, (c)\kmkl: the kernel is a combination of all base kernels using EasyMKL.
For each baseline, the computation of all kernels is required.
The hyperparameters are chosen using a nested 10-fold CV procedure, and they are $C\in\{2^i, i:-3\dots6\}$, and $\Lambda\in\{0,0.1\dots1\}$ for \kmkl.

@@

To learn $\hat{\boldsymbol{\eta}}$ we chose a random subsampling of training data and perform a 10-fold cross validation using EasyMKL to combine kernels and SVM as base learner. Then we fit the model with the best combination and select the top kernels.
In the second step we calculate the top kernel matrices using the whole training data and combine it to get the kernel. We propose 3 different methods to perform this combination:
(1)\ksmkl: using the EasyMKL algorithm, (2)\kssum: calculating the average of kernels, (3)\ksw: using the same weights (rescaled) from the first learning step.
In all of these methods we use a SVM as base learner. The hyperparameters sets are the same used in baselines.\\
\tablename\ \ref{tab:res} shows the AUC score, standard deviation and computation time used to make kernels and models for all problems.

>>>>

Model selection is one of the most computationally expensive tasks in a
machine learning application. When dealing with kernel methods for structures,
the choice with the largest impact on the overall performance is the selection
of the feature bias, i.e. the choice of the concrete kernel for structures.
Each kernel in turn exposes several hyper-parameters which also need to be
fine tuned. Finally, the downstream learning algorithm has its own hyper-
parameters. A common approach to hyper-parameters optimization is to fix
a-priori  a \textit{parameters grid}, i.e. a set of possible values for each
parameter. Then, the predictive performance of the learning algorithm is
evaluated for each possible parameters combination ({\em grid-search}
strategy), which is possibly a very time consuming step.  A more efficient
approach has been proposed in~\cite{FabioAiolliMicheleDonini2015,Massimo2016},
where instead of selecting the best parameters configuration, Multiple Kernel
Learning (MKL) was used for combining different kernel matrices in a single
learning procedure. While more efficient, this approach does not scale to very
large datset since a kernel matrix has to be computed for each kernel and for
each of its hyper-parameters. In this paper we propose a new approach that,
based on sub-sampling and MKL, is able to speedup the parameter selection
phase by orders of magnitude on real-world datasets  without loss in
predictive performance.

@@

In~\cite{Costa2010} the Neighborhood Subgraph Pairwise Distance Kernel (NSPDK) is presented. This kernel computes the exact matches between pairs of subgraphs with controlled size and distance (hyper-parameters $r$ and $d$, respectively). 
%The kernel has $O(|V| |V_h| |E_h| \log |E_h|)$ time complexity, where $|V_h|$ and $|E_h|$ are the number of nodes and the number of edges of the subgraph obtained by a breadth-fist visit of depth $h$. 
% The complexity of the kernel is $O(|V| |V_h| |E_h| \log |E_h|)$, where $|V_h|$ and $|E_h|$ are the number of nodes and the number of edges of the subgraph obtained by a breadth-fist visit of depth $h$. 
%The authors state that, for small values of the subgraph size and distance, the complexity of the kernel becomes in practice linear.

@@

\section{Experimental Assessment}
We conducted our experiments on several real-world binary datasets, that are MUTAG, CPDB, GDD, AIDS, NCI1, NCI109, CAS (see \cite{DaSanMartino2016,Shervashidze2011} for a more detailed description).
%AIDS~\cite{Weislow1989a}, CAS\footnote{http://cheminformatics.org/datasets/bursi/},
%CPDB~\cite{Helma2004a}, NCI1~\cite{Wale2006} and GDD~\cite{dobson2003}.
%The first four datasets contain chemical particles encoded in graph form, while the latter encodes proteins.
%{\color{red} The AIDS dataset contains chemical compounds (encoded in graph form), labeled according to their antiviral activity; CAS and CPDB store compounds labeled according to their mutagenicity; NCI1 and NCI109 consist of chemical compounds screened for activity against non-small lung cancer cells; GDD is composed of X-ray crystal structures of proteins; MUTAG contains %MUTAG serve per far vedere che funziona bene anche su dataset piccolissimi
%[...].
%All the nodes are labeled and there are no self loops.}
%
The kernels used in our experiments are:  NSPDK  with $h \in \{0,1\dots9\}$ and $d \in \{0,1\dots9\}$,  ODD$_{ST}$ with $h \in \{0,1\dots9\}$ and $\lambda \in \{0.5, 0.7,0.8\dots1.2,1.4,1.6,1.8\}$, WL with $h \in \{0,1\dots9\}$.
In all models, we considered kernels calculated in less that 1 hour, rejecting others.
The baselines used to compare our method consist on a SVM with kernels (a)\kval: the kernel is the one which has the best validation score, (b)\ksum: the kernel is the average of all base kernels, (c)\kmkl: the kernel is a combination of all base kernels using EasyMKL.
For each baseline, the computation of all kernels is required.
The hyper-parameters are chosen using a nested 10-fold CV procedure, and they are $C\in\{2^i, i:-3\dots6\}$, and $\Lambda\in\{0,0.1\dots1\}$ for \kmkl.

@@

To learn $\hat{\boldsymbol{\eta}}$ we chose a random subsampling of training data and perform a 10-fold cross validation using EasyMKL to combine kernels and SVM as base learner. Then we fit the model with the best combination and select the top kernels.
In the second step we calculate the top kernel matrices using the whole training data and combine it to get the kernel. We propose 3 different methods to perform this combination:
(1)\ksmkl: using the EasyMKL algorithm, (2)\kssum: calculating the average of kernels, (3)\ksw: using the same weights (rescaled) from the first learning step.
In all of these methods we use a SVM as base learner. The hyper-parameters sets are the same used in baselines.\\
\tablename\ \ref{tab:res} shows the AUC score, standard deviation and computation time used to make kernels and models for all problems.
